---
title: "Part 2 - Chi-Square My Version"
author: "Hannah Brady"
date: "12/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
```

## Read in data

```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```

Viewed the data for trapped lizards in Hornada Basin LTER.

We are going to focus on site and tail-breakage and proportions!

## Get counts of lizards by species

Species is spp. -n arranges counts in decreasing order.

```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n)
```

UTST species has most counts so we will just investigate that species.

## Make a subset

- UTST lizards
- Sites: IBPE (enclosure grassland), RABB (rabbit mesquite site), EAST (east tarbrush site)
- Tail condition is recorded (W = whole, B = broken)

Are proportion of lizards with broken/whole tails the same across sites?

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("B", "W"))
```

## Find counts and proportions of lizards with broken or whole tails at each site

A few ways to get the same info but one way that gets it in a prettier table format for chi-square.

Use count function from dplyer and count by site and tail:

```{r}
utst %>% 
  count(site, tail)
```

Provides table in long format, not contingency table which is what we need for chi-square.

Could use pivot-wider to turn the tail column into a B and W column.

OR

Use easier function from janitor package.

```{r}
tail_counts <- utst %>% 
  janitor::tabyl(site, tail)
tail_counts
```

Produces nice format that is almost contingency table format.

But we also want to present the counts with proportions directly.

```{r}
tail_proportions <- tail_counts %>% 
  # Add proportions:
  adorn_percentages() %>% 
  # Round the digits:
  adorn_pct_formatting(digits = 2) %>% 
  # Add the counts back in:
  adorn_ns()
# Looks so pretty! Just throw it in kable for final styling.
```

Gives proportions by site. Can see that EAST and IBPE have very similar proportions of B to W but that RABB is different. Use chi-square to figure out if it's significantly different.

## COnvert to a contingency table

Formatting for chi-square test contingency table:

Cannot have the names of the sites be interpreted by R as a column of data. Convert it to "row names" format so it is not read as data. Site is no longer considered a variable. Use `column_to_rowname()`

```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```

Once only thing existing in your table is counts, you can run it through the chi-square test.

## Chi-square

Are tail breakage and site independent of each other?
Recall there are like three ways to ask this question that all mean the same thing.

```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

lizard_tidy <- broom::tidy(lizard_x2)
```

Outcome interpretation:

chi-squared test statitic value is an aggregate indicating how diff our values are from expected null hypothesis values.

See that df = 2:

$$degrees of freedom = (rows - 1)*(columns - 1) = (3-1)*(2-1) = 2$$

If you have a chi-squared value of 13 with 2 degrees of freedom then you get a p-value of 0.0015.

Is that probability so low that I can reject the null hypothesis? Yes. Retain alternative hypothesis: site and tail breakage are NOT indepedent.

Also: There is a significant association between site and tail breakage. There is a significant effect of site on tail breakage. Site and lizard tail breakage are not independent.

But make these phrases reproducible:

There is a significant association (i.e. non-independence) between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter` = `r round(lizard_tidy$statistic,2)`, *p* = `r round(lizard_tidy$p.value,3)`)).
