---
title: 'Week 9, Lab Part 1: MLR cont.'
author: "Hannah Brady"
date: "11/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(GGally)
library(stargazer)
```

## Step 1: Attach packages and look at penguins

In console did View(penguins).

## Step 2: Reminder visual data exploration

Use ggpairs to look at relationships of all variables.

```{r}
penguins %>% 
  ggpairs()
# whoa that's overwhelming
```

Parse it down to make it easier to understand.

```{r}
penguins %>% 
  select(species, bill_length_mm:body_mass_g) %>% 
  ggpairs(aes(color = species))
```

Within groups between my different quantitative variables, it looks like a linear relationship makes sense.

### Determining non-collinearity
Scatterplots show that nothing is too highly correlated. The overall coreelation is shown to be 0.871

Highest correlation is Gentoo body_mass_g by bill_depth_mm at 0.719. Still not crazy high. Nothing is highly correlated, would not be overfitting model to add any variable based on collinearity.

Seems like multiple linear regression is a good option.

## Build a few different models

```{r}
lm1 <- lm(body_mass_g ~ flipper_length_mm + species, data = penguins)
lm1
```

Adelie is reference level for species.
40.71 mass increase for each unit of flipper length increase.
If all other variables are equal, a Chinstrap penguin will have 206.51 times less mass than an Adelie.

```{r}
lm2 <- lm(body_mass_g ~ flipper_length_mm + species + sex, data = penguins)
lm2

lm3 <- lm(body_mass_g ~ flipper_length_mm + species + sex + bill_length_mm, data = penguins)
lm3

lm4 <- lm(body_mass_g ~ flipper_length_mm + species + sex + bill_length_mm + island, data = penguins)
lm4
```

```{r}
 plot(lm1) # No concerns
 plot(lm2) # No concerns
 plot(lm3) # No concerns
 plot(lm4) # No concerns
```

AIC

Penalizes addition of more variables and lets you know if improved model fit was worth it.

Lower the value of the AIC - the better.

# Find AIC for all models

No packages needed, part of base R's stats package

```{r}
AIC(lm1)
# Value by itself, means nothing! Only useful when comparing permutations of a model.
AIC(lm2)
AIC(lm3)
AIC(lm4)
```

AIC loved adding sex.
Didn't like addition of island in lm4 - not worth it!
If only using AIC value as metric - lowest AIC value was for lm3, indicating that is the best balance of model fit and model complexity.

Recall that it is only part tho of how we evaluate and make models. Think about what we already know, bg research, conceptually.

## Use stargazer for a table with multiple model outputs

```{r, results = 'asis'}
stargazer(lm1, lm2, lm3, lm4, type = "html")
```

Nice table!
Helps you see how coefficients change as you add variables.
Great way to get to almost publication ready tables with multiple models.

Example coefficient interpretation:
lm3 bill_length_mm:
1 mm increase in bill length is associated with a 21.633 g increase in body mass.
Parenthesis below are the standard error.

R-squared is similar, so if we were only using models for predictive capacity might not need to add variables.

## Omitted variable bias

```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, y = bill_depth_mm))+
  geom_point() +
  geom_smooth(method = "lm")
```

```{r}
ggplot(data = penguins, aes(x = flipper_length_mm,
                            y = bill_depth_mm,
                            group = species))+
  geom_point(aes(color = species)) +
  geom_smooth(method = "lm")
```

Species completely changes the correlation. Note actual points didn't change
